---
title: "ST558 Project 2  Interacting with APIs: Example with the Financial API"
author: "Xi Zeng"
date: "2022-10-06"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction  
This document is a vignette to show how to retrieve data from an API. In this project, we will be interacting with the API for financial data. Also, we will build a few functions to interact with some of the endpoints as well as retrieve data to perform a basic exploratory data analysis.  

# Requirements  
To use the functions for interacting with the Financial API, here are some packages needed:  
- `tidyverse`: Collections of packages with useful functions for data manipulation and visualization  
- `jsonlite`: Package needed to interact with API and parse data  
- `httr`: Used for load the data we need from API  

# API Interaction Functions  
Here is the part for the self-defined functions to interact with the Financial API, as well as some helper functions.  

`Function1`  

This function is for interacting with the `Daily Open/Close` endpoint for the Financial API.  

```{r}
library(tidyverse)
library(httr)
library(jsonlite)

Mydata <- GET("https://api.polygon.io/v2/aggs/ticker/AAPL/range/1/day/2020-06-01/2020-06-17?apiKey=pxkABpfvwGyIl55tOWtNxricS6IwKFYH")
str(Mydata)
Mydata$content
```



## Including Plots

You can also embed plots, for example:


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
```{r, eval = FALSE}
library(rmarkdown)
rmarkdown::render(input = "ST558_update_check.Rmd",
                  output_file = "README.md")
```
